---
title: "2017-02-13"
author: "Rick O. Gilmore"
date: "`r Sys.Date()`"
bibliography: ../bib/reproducibility.bib
output:
  ioslides_presentation:
    css: ../css/gilmore-ioslides.css
    widescreen: true
    incremental: false
    transition: default
    self-contained: true
---

<!-- Lecture -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Today's topics

- Simulation as a tool for reproducible and transparent science
- Visualization tools in R

## Simulation as a tool for reproducible and transparent science

- Why simulate
- What to simulate
- How to simulate

## Why & what to simulate?

- Explore sample sizes, effect sizes, power
- Pre-plan/test, polish data-munging workflows
- Make hypotheses even more explicit
    + Simulation == Pregistration on steroids
    + '~~X affects Y~~' -> 'Mean(X) > Mean(Y)'
    + or 'Mean(X) >= 2*Mean(Y)'
- Simulate data analysis in advance

---

- Plan data visualizations in advance
- Avoid avoidable errors
- Plan your work, work your plan
- Super easy to run analyses when your data come in

## How to simulate

- R functions
- R Markdown document(s)

## Super-simple example

- Hypothesis 1: Height (inches) is correlated with weight (lbs)


```{r pick-parameters}

# choose sample size
sample.n <- 200

# choose intercept and slope
beta0 <- 36   # inches
beta1 <- 0.33  # Rick's guess

# choose standard deviation for error
sigma <- 10 # Rick's guess
```

---

```{r generate-data}
# random weights between 80 lbs and 250 lbs (uniform sampling)
w <- runif(n = sample.n, min = 80, max = 250)

h.pred <- rep(x = beta0, n = sample.n) + beta1 * w
h <- h.pred + rnorm(n = sample.n, mean = 0, sd = sigma)
```

<!-- Code to make scrolling slides -->
<style>
slides > slide { overflow: scroll; }
slides > slide:not(.nobackground):before {
  background: none;
  }
slides > slide:not(.nobackground):after {
  content: '';
  background: none;
  }
}
</style>
---

```{r plot-hist}
library(ggplot2)
library(dplyr)

hist(w)
hist(h)
hist(h.pred)
```

---

```{r plot-scatter}
# Put h and w into data frame for ggplot
height.weight <- data.frame(inches = h, lbs = w)

# Plot
scatter.1 <- ggplot(data = height.weight) +
  aes(x = lbs, y = inches) +
  geom_point()
scatter.1
```

## That's synthesis, now analysis

- Remember Hypothesis 1: Height (inches) is correlated with weight (lbs)?

```{r measure-correlation}

# Could use the raw data
# cor.test(x = w, y = h)
# Or, to use the values in the data frame, use with(...)

with(height.weight, cor.test(x = inches, y = lbs))
```

## Aside: extracting the statistics to make an interactive report

```{r save.cor.stats}
# Save output as a variable
cor.test.inches.lbs <- with(height.weight, cor.test(x = inches, y = lbs))

# What sort of beast is this?
mode(cor.test.inches.lbs)
```

---

```{r unpack-list}
# Aha, it's a list, this shows me all of the parts
unlist(cor.test.inches.lbs)

# Looks like the t value is the first element
cor.test.inches.lbs[[1]]
```

---

The `r cor.test.inches.lbs[[7]]` correlation between height and weight is `r cor.test.inches.lbs[[4]]`, t(`r cor.test.inches.lbs[[2]]`)=`r cor.test.inches.lbs[[1]]`, p=`r cor.test.inches.lbs[[3]]`, with a 95% confidence interval of [`r cor.test.inches.lbs[[9]][1]`, `r cor.test.inches.lbs[[9]][2]`].

<span class="blue">Obviously, we should do some formatting before submitting this, but you get the idea.</span>

## Now back to analysis with our synthetic data

```{r linear-reg}
fit <- lm(formula = inches ~ lbs, data = height.weight)
summary(fit)
(ci <- confint(fit)) # saves in variable ci and prints
```

## How'd we do?

| Parameter | Actual    | Low Estimate | High Estimate |
|-----------|-----------|--------------|---------------|
| $\beta0$  | `r beta0` | `r ci[1,1]`  | `r ci[1,2]`   |
| $\beta1$  | `r beta1` | `r ci[2,1]`  | `r ci[2,2]`   |

- Why off on the slope ($\beta1$)
- Random error, probably. Could run again.

---

```{r run-it-all-again, echo=FALSE}
# random weights between 80 lbs and 250 lbs (uniform sampling)
w <- runif(n = sample.n, min = 80, max = 250)

h.pred <- rep(x = beta0, n = sample.n) + beta1 * w
h <- h.pred + rnorm(n = sample.n, mean = 0, sd = sigma)
height.weight <- data.frame(inches = h, lbs = w)

fit <- lm(formula = inches ~ lbs, data = height.weight)
summary(fit)
(ci <- confint(fit)) # saves in variable ci and prints
```

---

| Parameter | Actual    | Low Estimate | High Estimate |
|-----------|-----------|--------------|---------------|
| $\beta0$  | `r beta0` | `r ci[1,1]`  | `r ci[1,2]`   |
| $\beta1$  | `r beta1` | `r ci[2,1]`  | `r ci[2,2]`   |

# Visualization in R

## Plot first, analyze last

>- Why?
  + [Mike Meyer](https://www.linkedin.com/in/mike-meyer-0056578a) told me so
  + Less biased
  + Easier to be transparent and reproducible
  + Want/need to plot eventually anyway
  + If a picture's worth a thousand words...
>- How?

## How

- Base graphics
    + `plot(x,y)` `hist(x)`, `coplot()`
- `ggplot2`
    + Grammar of graphics
    
## Base graphics

- Try it, maybe you'll like it
- `plot()` takes many types of input
- So does `summary()`
- A little harder to customize

## Data visualization with ggplot2

Wickham, H. & Grolemund, G. (2017). *R for Data Science*. O'Reilly. <http://r4ds.had.co.nz/>

## Let's just walk through Chapter 3

<http://r4ds.had.co.nz/data-visualisation.html>

## Other ggplot2 resources

- Wickham, H. (2010). *ggplot2: Elegant Graphics for Data Analysis (Use R!)* <http://ggplot2.org/book/>
- ggplot2 2.1.0 documentation: <http://docs.ggplot2.org/current/>
