---
title: "2017-01-16-workflows"
author: "Rick O. Gilmore"
date: "`r Sys.time()`"
bibliography: ../bib/reproducibility.bib
always_allow_html: yes
output:
  ioslides_presentation:
    css: ../css/gilmore-ioslides.css
    widescreen: true
    incremental: false
    transition: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Today's Topics

- Workflows and **methods** reproducibility [[@goodman_what_2016]](http://doi.org/10.1126/scitranslmed.aaf5027)

## [Workflows](https://en.wikipedia.org/wiki/Workflow)

<"https://www.google.com/search?q=workflow">

## Typical workflows in experimental psychology

- Idea/question/hypotheses
- Design study
- Seek ethics board permission
- Build/borrow/buy data collection instruments
- Run study
- Analyze results
- Write-up results for presentation and/or publication

## Design study

- Participants
    + Number, characteristics
- Setting(s)
    + Lab, classroom, home
- Measures or tasks
    + Self/other report
    + Observations/video or audio recording
    + Physiological measures (MRI, EEG, ECG)
    + Computer-based tasks

## Data collection instruments

- Surveys
- Video/audio
- MRI, EEG, ECG
- Computer-generated data files

## Run study

```
done.collecting.data = FALSE
while (!done.collecting.data) {
  Collect.sample()
  if (collected.sample.n >= planned.sample.n) {
    done.collecting.data = TRUE
  } else {
    done.collecting.data = FALSE
  }
}
```

## Analyze results 

- Clean/check data
- Merge, combine, munge data

----

<div class="centered">
```{r, echo=FALSE}
library(DiagrammeR)
grViz("
      digraph {
      layout = dot
      node [shape = square]
      Analyze -> Evaluate
      Evaluate -> Analyze
      }")
```
</div>

## Prepare presentation/publication

- Intro
- Methods
- Results
    + Stats
    + Plots
- Conclusions
- References
- Data?

## Behavioral study summary

<div class="centered">
```{r, echo=FALSE}
grViz("
      digraph {
      layout = dot
      node [shape = circle]
      E1 -> XL1
      S1 -> XL1
      E2 -> XL2
      S2 -> XL2
      En -> XLn
      Sn -> XLn
      {XL1, XL2, XLn} -> XLall
      XLall -> Stat
      Stat -> {XLplot, MSWd, Ppt}
      XLplot -> {MSWd, Ppt}
      }")
```
</div>
----
<div class="centered">

### Imaging example
```{r, echo=FALSE}
grViz("
      digraph {
      layout = dot
      node [shape = circle]
      {Str1, MNI} -> Norm1
      {Norm1, F1} -> Fnorm1
      {Str2, MNI} -> Norm2
      {Norm2, F2} -> Fnorm2
      {StrN, MNI} -> NormN
      {NormN, Fn} -> FnormN
      {Fnorm1, Des1, Beh1} -> FstL1
      {Fnorm2, Des2, Beh2} -> FstL2
      {FnormN, DesN, BehN} -> FstLn
      {FstL1, FstL2, FstLn} -> SndL
      {SndL, Beh} -> {Plot, MSWd, Ppt}
      Plot -> {MSWd, Ppt}
      }")
```
</div>

## Threats to **methods** reproducibility

- Idea/question/hypotheses
- Design study
- Seek ethics board permission
- Build/borrow/buy **data collection instruments**
- **Run study**
- **Analyze results**
- **Write-up results** for presentation and/or publication

## What are the threats?

- Data collection instruments
- Running study
- Data analysis
- Study write-up

## Mitigating the threats

- Maximize consistency, **methods** reproducibility [[@goodman_what_2016]](http://doi.org/10.1126/scitranslmed.aaf5027)
- Design of/consistent adherence to detailed experimental protocols
- Consistent, transparent workflows
- Consistent, transparent organization of data, metadata
- Minimize human/hand data entry
- Automate as much as possible

## How detailed is your (internal) protocol?

- Play & Learning Across a Year (PLAY) project [wiki](https://dev1.ed-projects.nyu.edu/wikis/docuwiki/doku.php/collection)

## Questions to consider

- What data and metadata am I collecting?
- How does it get collected?
- Where does it go after it's collected?
- How does my non-electronic data get transferred to an electronic form?
- How do my electronic data files get cleaned, merged, munged?

## Reproducible workflow aspirations

- "Chain of custody" from raw data to finished results and figures
- Single command to regenerate all results and figures from raw data

<http://datasci.kitzes.com/lessons/python/reproducible_workflow.html>

## Reproducible workflow recommendations

- Create consistent structure for projects
    + Use file name conventions
- Use machine-readable file types
    + commas-separated value (.csv) vs. .xlsx
- Automate as much as possible
- Use version control

## Lots of ways to organize electronic data...

```
study-1/
  sub-001/
    sub-001-measure-a.txt
    sub-001-image.jpg
    sub-001-demo.csv
    sub-001-measure-b.txt
  sub-002/
    sub-002-measure-a.txt
    sub-002-image.jpg
    sub-002-demo.csv
    sub-002-measure-b.txt
  ...
  sub-00n/
    ...
```

----

```
study-1/
  measure-a/
    sub-001-measure-a.txt
    ...
  measure-b/
    sub-001-measure-b.txt
    ...
  image/
    sub-001-image.jpg
    sub-002-image.jpg
    ...
  demo/
    sub-001-demo.csv
    sub-002-demo.csv
    ...
```

----

```
study-1/
  analysis/
    data/
      sessions/
        2017-01-09-sub-001/
        ...
      aggregate/
        study-1-demo-aggregate.csv
        study-1-measure-a-aggregate.csv
        ...
    R/
    img/
    reports/
  protocol/
    code/
      my-experiment.m
    materials/
      stim-1.jpg
      stim-2.jpg
      ...
```

----

```
  pubs/
    presentations/
    papers/
  refs/
  grants/
    2016/
    2017/
  irb/
  mtgs/
```

## Databrary's volume, session/materials model

<iframe src="https://nyu.databrary.org/volume/2">
  <p>Your browser does not support iframes.</p>
</iframe>

<https://nyu.databrary.org/volume/2>

## [ProjectTemplate](http://projecttemplate.net/index.html)

- Automates some of the project management involved in data analysis
    + Hat Tip (HT): Michael Hallquist
- Gilmore says: Use what you like

## Can automate project creation, too

```
## Create project directory
proj.name = "tmp_proj"
if (!exists(proj.name)) {
  dir.create(path = proj.name, recursive = TRUE)
}

# Create sessions directory
sessions.dir = paste(proj.name, "analysis/sessions", sep="/")
if (!exists(sessions.dir)) {
  dir.create(path = sessions.dir, recursive = TRUE) # creates intermediate dirs
}

# Aggregate data file directory
agg.dir = paste(proj.name, "analysis/aggregate", sep="/")
if (!exists(agg.dir)) {
  dir.create(path = agg.dir, recursive = TRUE)
}
```

## Words to the wise

- Use consistent file/directory names
    + lowerCamelCaseIsGood.txt so is UpperCamelCase.txt
    + underscores_between_words.txt works; so do dashes-between.txt
    + avoid spaces in your file names.txt; these are not always reliably readable by all computers.
- Choose good, [descriptive names](http://kbroman.org/dataorg/pages/names.html)

## Consider seriously [Karl Broman's guides](http://kbroman.org/dataorg/)

- Be consistent
- Write dates as YYYY-MM-DD.
- Fill-in all cells
- One thing in a cell
- Make your data a rectangle
- Create a data dictionary

## Consider seriously [Karl Broman's guides](http://kbroman.org/dataorg/)

- No calculations in raw data files
- No font or color to highlight data
- Make back-ups
- Validate data to avoid data entry errors
- Save data in plain text files (comma or tab-delimited)

## Why?

- Data scientists (that's you!) spend a lot of time just cleaning data
- <http://www.infoworld.com/article/3047584/big-data/hottest-job-data-scientists-say-theyre-still-mostly-digital-janitors.html>

## Easy to merge data sets if they contain a linking variable (like subID)

- [study-1-demo-agg.csv](study-1-demo-agg.csv) contains
    + subID, sex, ageYrs, favColor
- [study-1-rt-agg.csv](study-1-rt-agg.csv) contains
    + subID, condition, rt
    
----

```
subID,sex,ageYrs,favColor
001,m,53,green
002,f,51,blue
003,f,23,red
004,m,25,aqua
```

Don't put spaces between variables in comma-separated value (.csv) files.
Also, make sure to add a final line feed/enter character.

----

```
subID,condition,rt
001,upright,250
001,inverted,300
002,upright,225
002,inverted,290
003,upright,270
003,inverted,230
004,upright,210
004,inverted,240
```
    
----
```{r, include=TRUE, echo=TRUE}
# read data files, first row (header) contains variable names
demo <- read.csv(file = "study-1-demo-agg.csv", header = TRUE)
rt <- read.csv(file = "study-1-rt-agg.csv", header = TRUE)

# merge and print
merged <- merge(demo, rt, by = "subID")
merged
```

## A final word about tidy data [[@hadley_wickham_tidy_2014]](http://doi.org/10.18637/jss.v059.i10)

- Variables in columns
- Observations in rows
- ~~Ok~~/better to repeat values in columns
    + subID,trial,rt
    + 001,1,300
    + 002,2,327
    + 003,3,429

## Main points

- Think like a computer!
- Plan your work; work your plan
- Consistency, standard formats
- Tidy data
- Haven't said anything about "openness"...yet

## More resources

- Data Carpentry, <http://www.datacarpentry.org/>
- Software Carpentry, <https://software-carpentry.org/>
- Open Science Framework (OSF), <http://osf.io>
- R for Data Science, <http://r4ds.had.co.nz/>

## References {.smaller}
