---
title: '2017-02-13'
author: "Rick O. Gilmore"
date: '`r Sys.Date()`'
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: 2
  ioslides_presentation:
    css: ../css/gilmore-ioslides.css
    incremental: no
    self-contained: yes
    transition: default
    widescreen: yes
  html_document:
    keep_md: yes
    self_contained: yes
bibliography: ../bib/reproducibility.bib
---
<!-- In-class activities -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

list.of.packages <- c("ggplot2", "DescTools")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Loading packages
lapply(list.of.packages, require, character.only = TRUE)
```

## Goals

- Advanced simulation
- Practice with visualization

## Advanced simulation

- Putting code into a function
- A factorial design?

## Let's make it a function

- Functions help automate routines
- Parts of a function:
    + Input parameters
        - Defaults or not
    + Output(s)
- `My.function.name <- function(my.param1, my.param2 = "cool")`

----

```{r define-global-constants}

# define global constants from prior simulation
sample.n = 200
beta0 = 36
beta1 = .33
sigma = 10
min.x = 80
max.x = 250
```


<!-- Code to make slides that overflow scroll. -->
<!-- http://stackoverflow.com/q/38260799 -->
<style>
slides > slide { overflow: scroll; }
slides > slide:not(.nobackground):before {
  background: none;
  }
slides > slide:not(.nobackground):after {
  content: '';
  background: none;
  }
}
</style>

---

```{r define-function}
Height.weight.sim <- function(sample.n = 200, beta0 = 36, beta1 = .33, sigma = 10, min.x = 80, max.x = 250) {
  # Calculates correlation, intercept, slope estimates for
  # linear relation between two variables
  
  # Args:
  #   sample.n: Number of sample poings, default is 200
  #   beta0: Intercept, default is 36 (inches)
  #   beta1: Slope, default is .33
  #   sigma: Standard deviation of error
  #   min.x: Minimum value for x (weight in lbs)
  #   max.x: Maximum value for x (weight in lbs)
  #
  # Returns:
  #   Named array with values
  #   beta0
  #   beta1
  #   beta0.lo: 2.5% quantile for intercept
  #   beta0.hi  97.5% quantile for intercept
  #   beta1.lo  2.5% quantile for slope
  #   beta1.hi  97.5% quantile for slope

  w <- runif(n = sample.n, min = min.x, max = max.x)

  h.pred <- rep(x = beta0, n = sample.n) + beta1 * w
  h <- h.pred + rnorm(n = sample.n, mean = 0, sd = sigma)
  height.weight <- data.frame(inches = h, lbs = w)

  fit <- lm(formula = inches ~ lbs, data = height.weight)
  ci <- confint(fit)
  
  # Create output vector with named values
  (results <- c("beta0" = beta0, 
    "beta1"= beta1, 
    "beta0.lo" = ci[1,1],
    "beta0.hi" = ci[1,2],
    "beta1.lo" = ci[2,1],
    "beta1.hi" = ci[2,2]))
}
```

---

```{r run-height-weight-function}
# Defaults only
Height.weight.sim()

# Larger sample size
Height.weight.sim(sample.n = 500)
```

## Doing a series of simulations

- Goal: run our function a number of times, collect the results

```{r run-n-simulations}
n.simulations = 100
n.vars = 6 # variables Height.weight.sim() outputs

# initialize output array
height.weight.sim.data <- array(0, dim=c(n.simulations, n.vars))

# Repeat Height.weight.sim() n.simulations times
for (i in 1:n.simulations) { 
  height.weight.sim.data[i,] <- Height.weight.sim()
  }
```

---

```{r plot-n-simulation-results}

# Easier to make simulation data a data frame
ht.wt.sims <- as.data.frame(sim.num = 1:n.simulations, height.weight.sim.data)

# Add a variable to index the simulation number
ht.wt.sims$sim.num <- 1:n.simulations

# Plot beta0 min and max (now called V3, V4)
ggplot(data = ht.wt.sims) +
  aes(x = sim.num) +
  geom_linerange(mapping = aes(ymin=V3,
                                ymax=V4)) +
  ylab("CIs for Intercept") +
  geom_hline(yintercept = beta0, color = "red")
```

---

```{r}
# Plot beta1 min and max (now called V5, V6)
ggplot(data = ht.wt.sims) +
  aes(x = sim.num) +
  geom_linerange(mapping = aes(ymin=V5,
                                ymax=V6)) +
  ylab("CIs for Slope") +
  geom_hline(yintercept = beta1, color = "red")
```

## How about a factorial design?

- Dependent variable: RT
- Independent variables
    + Fixed
        - Symbol type: {letter, number}
    + Random
        - Subject mean RT
- Hypothesis:
    + RT to detect numbers is lower than for letters
    + There is no order effect
- Rick wishes he'd found this first: <https://www.r-bloggers.com/design-of-experiments-%E2%80%93-full-factorial-designs/>

---
```{r factorial-design}
# Simulation parameters
n.subs = 30
trials.per.cond = 100

letters.numbers.rt.diff = 50 # ms
rt.mean.across.subs = 35
sigma = 50
cond.labels = c("letter", "number")
cond.rts <- c("letter" = 0, "number" = letters.numbers.rt.diff)

stim.types <- factor(x = rep(x = c(1,2), trials.per.cond), labels = cond.labels)

#sample(factor(x=rep(c("letter", "number"), 100)), 200)
random.stim.types <- sample(stim.types, trials.per.cond*length(cond.labels))

mean.sub.rt <- rnorm(n = 1, mean = rt.mean.across.subs, sd = sigma)
trial.rt <- array(0, dim = length(random.stim.types))

# Generate RTs based on trial, condition
for (t in 1:length(random.stim.types)) {
  trial.rt[t] <- mean.sub.rt + cond.rts[random.stim.types[t]] + rnorm(n = 1, mean = 0, sd = sigma)
}

# Make data frame
letter.number.df <- data.frame(trial = 1:length(random.stim.types), stim = random.stim.types, rt = trial.rt)

ggplot(data = letter.number.df, aes(x = random.stim.types, y = rt)) + geom_boxplot()
```

## Put this in a function

```{r make-factorial-function}
Simulate.sub.rt <- function(trials.per.cond = 100,
                            letters.numbers.rt.diff = 50,
                            rt.mean.across.subs = 350, 
                            sigma = 50) {
  
  cond.rts <- c("letter" = 0, "number" = letters.numbers.rt.diff)
  stim.types <- factor(x = rep(x = c(1,2), trials.per.cond), labels = c("letter", "number"))
  random.stim.types <- sample(stim.types, 200)

  mean.sub.rt <- rnorm(n = 1, mean = rt.mean.across.subs, sd = sigma)
  trial.rt <- array(0, dim = length(random.stim.types))

  # Generate RTs based on trial, condition
  for (t in 1:length(random.stim.types)) {
    trial.rt[t] <- mean.sub.rt + cond.rts[random.stim.types[t]] +
      rnorm(n = 1, mean = 0, sd = sigma)
    }

  # Make data frame
  letter.number.df <- data.frame(trial = 1:length(random.stim.types), 
                                 stim = random.stim.types, 
                                 rt = trial.rt)
}
```

## Create code to a set of participants and create one merged data frame

```{r make-merged-simulation-df}
Make.sub.rt.df <- function(sub.id) {
  sub.rt.df <- Simulate.sub.rt()
  sub.rt.df$sub.id <- sub.id
  sub.rt.df
}

# Use lapply to make separate data frames for all subs
sub.rt.df.list <- lapply(1:n.subs, Make.sub.rt.df)

# Use Reduce() with the merge function to make one big file
sub.rt.df.merged <- Reduce(function(x, y) merge(x, y, all=TRUE), sub.rt.df.list)
```

## Now, want to see what we have?

```{r plot-merged-simulation-data}
ggplot(data = sub.rt.df.merged) +
  aes(x=stim, y=rt, color=stim) +
  geom_boxplot() +
  facet_grid(facets = . ~ as.factor(sub.id))
```

## And, just for fun

```{r one-way-anova}
library(lme4)
fit1 <- lmer(formula = rt ~ stim + (1|sub.id), data = sub.rt.df.merged)
summary(fit1)
```

## Or

```{r one-way-using-aov}
fit2 <- aov(formula = rt ~ stim + Error(as.factor(sub.id)), data = sub.rt.df.merged)
summary(fit2)
```

## Practice with visualization

- Work through the Wickham & Grolemund chapter on your own.
    + <http://r4ds.had.co.nz/data-visualisation.html>

## Learning by example

```
# Installing packages
list.of.packages <- c("xlsx", "ggplot2","nlme", "foreign", "gplots", "stats", "psych" , "ISwR", "ggm" , "car" , "ppcor", "StatDA", "plyr","DescTools")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Loading packages
lapply(list.of.packages, require, character.only = TRUE)
```
<https://github.com/psu-psych-511-2017-spring/liu_yushuang/blob/master/output.Rmd>

## A syntax I like

```
my.data <- data.frame(x.var = ..., y.var = ...)

my.data %>% 
  ggplot() +      # pipe %>% operator feeds it my.data
  aes(x=x.var, y=y.var) + # plus + operator 'adds' to plot
  geom_point() -> # assign -> operator saves output 
  my.scatter.plot # variable containing plot
  
my.scatter.plot # prints it
```
- Uses 'pipe' `%>%` operator and 'assign' `->` operator
- Easy to follow flow of what's going on

## Assignment (due Monday, Feburary 20)

- Create your own simulated data set for a real or proposed study.
    + You may adapt or build upon the examples used in class.
- Plot the results of your simulation using ggplot2 commands.
