---
title: "report-template"
author: "X. Ellent Student"
date: "`r Sys.time()`"
output: 
  html_document:
    self_contained: true
  word_document:
    fig_width: 5
    fig_height: 5
    fig_caption: false
    df_print: kable
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

This is a report template. 

## Introduction

If you hit the `Knit` button in RStudio, it will create an `html_document` because that is the first line in the YAML header. 

~~~
---
title: "report-template"
author: "X. Ellent Student"
date: "`r Sys.time()`"
output: 
  html_document:
    self_contained: false
  word_document:
    fig_width: 5
    fig_height: 5
    fig_caption: false
    df_print: kable
  pdf_document:
    toc: true
    toc_depth: 2
---
~~~

You can just render the document at the command line with `rmarkdown::render(input = "report-template.Rmd"`.

You can change the default output format by changing the order of the items in the header:

~~~
---
title: "report-template"
author: "X. Ellent Student"
date: "`r Sys.time()`"
output: 
  word_document:
    fig_width: 5
    fig_height: 5
    fig_caption: false
    df_print: kable
  html_document:
    self_contained: false
  pdf_document:
    toc: true
    toc_depth: 2
---
~~~

This renders a Word file by default.

To pick some but not all output formats, enter `rmarkdown::render(input = "report-template.Rmd", output_format = c("html_document", "pdf_document"))`

You can also go hog-wild and render all three output formats at once by entering `rmarkdown::render(input = "report-template.Rmd", output_format = "all")`.

## Methods

Describe what you did here.

### Participants

With whom...

### Materials

With what...

### Procedure

How...

### Analysis

What you measured...

```{r sample-data}
x = rnorm(n = 100, mean = 0, sd = 1)
y = rnorm(n = 100, mean = 5, sd = 10)
```

## Results

What you found...

### My scatterplot

```{r figure-1}
plot(x,y)
```

The correlation between x and y is `r cor(x,y)`.

### Some histograms

```{r figure-2}
hist(x)
hist(y)
```

The mean of x is `r mean(x)` and the standard deviation is `r sd(x)`. Or, try a table:

| Variable | Mean        | SD        |
|----------|-------------|-----------|
| x        | `r mean(x)` | `r sd(x)` |
| y        | `r mean(y)` | `r sd(y)` |

## Conclusion

What you concluded.