---
title: '2017-02-13'
author: "Rick O. Gilmore"
date: '`r Sys.Date()`'
output:
  ioslides_presentation:
    css: ../css/gilmore-ioslides.css
    incremental: no
    self-contained: yes
    transition: default
    widescreen: yes
  html_document:
    keep_md: yes
    self_contained: yes
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: 2
bibliography: ../bib/reproducibility.bib
---
<!-- Lecture -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Today's topics

- Some observations about GitHub use and R Markdown
- Simulation as a tool for reproducible and transparent science
- Visualization tools in R

## Observations about GitHub use

- You're in charge of what goes where.
- Public repos are public, but no one knows what you're doing unless you alert them.
    - If you create a repo in your own account, add `rogilmore` or other collaborators to use the `@rogilmore` type at-mentions feature.
    - Files you create in `https://github.com/psu-psych-511-2017-spring` I can already comment on.

## Observations about GitHub use

- Pull requests are when you edit my code and want me to "pull"/adopt it.
    - If I'm a collaborator on the project with `write` privileges, I don't have to issue a pull request.

## Observations about R Markdown

- Ok to make multiple R Markdown files
- Make sure to add spaces where they belong: `##Header` vs. `## Header`
- Comments! Add them. This is your record of what you did.
- Don't forget you can hide things

## Be a risk-taker; be your own professor

- <http://www.stat.cmu.edu/~cshalizi/rmarkdown>
- <http://stat545.com/bit006_github-browsability-wins.html>

## Simulation as a tool for reproducible and transparent science

- Why simulate
- What to simulate
- How to simulate

## Why & what to simulate?

- Explore sample sizes, effect sizes, power
- Pre-plan/test, polish data-munging workflows
- Make hypotheses even more explicit
    + Simulation == Pregistration on steroids
    + '~~X affects Y~~' -> 'Mean(X) > Mean(Y)'
    + or 'Mean(X) >= 2*Mean(Y)'
- Simulate data analysis in advance

---

- Plan data visualizations in advance
- Avoid avoidable errors
- Plan your work, work your plan
- Super easy to run analyses when your data come in

## How to simulate

- R functions
- R Markdown document(s)

## Super-simple example

- Hypothesis 1: Height (inches) is correlated with weight (lbs)

```{r pick-parameters}

# choose sample size
sample.n <- 200

# choose intercept and slope
beta0 <- 36   # inches
beta1 <- 0.33  # Rick's guess

# choose standard deviation for error
sigma <- 10 # Rick's guess
```

---

```{r generate-data}
# random weights between 80 lbs and 250 lbs (uniform sampling)
w <- runif(n = sample.n, min = 80, max = 250)

h.pred <- rep(x = beta0, n = sample.n) + beta1 * w
h <- h.pred + rnorm(n = sample.n, mean = 0, sd = sigma)
```

<!-- Code to make scrolling slides -->
<style>
slides > slide { overflow: scroll; }
slides > slide:not(.nobackground):before {
  background: none;
  }
slides > slide:not(.nobackground):after {
  content: '';
  background: none;
  }
}
</style>
---

```{r plot-hist}
library(ggplot2)
library(dplyr)

hist(w)
hist(h)
hist(h.pred)
```

---

```{r plot-scatter}
# Put h and w into data frame for ggplot
height.weight <- data.frame(inches = h, lbs = w)

# Plot
scatter.1 <- ggplot(data = height.weight) +
  aes(x = lbs, y = inches) +
  geom_point()
scatter.1
```

## That's synthesis, now analysis

- Remember Hypothesis 1: Height (inches) is correlated with weight (lbs)?

```{r measure-correlation}

# Could use the raw data
# cor.test(x = w, y = h)
# Or, to use the values in the data frame, use with(...)

with(height.weight, cor.test(x = inches, y = lbs))
```

## Aside: extracting the statistics to make an interactive report

```{r save.cor.stats}
# Save output as a variable
cor.test.inches.lbs <- with(height.weight, cor.test(x = inches, y = lbs))

# What sort of beast is this?
mode(cor.test.inches.lbs)
```

---

```{r unpack-list}
# Aha, it's a list, this shows me all of the parts
unlist(cor.test.inches.lbs)

# Looks like the t value is the first element
cor.test.inches.lbs[[1]]
```

---

The `r cor.test.inches.lbs[[7]]` correlation between height and weight is `r sprintf("%.3f", cor.test.inches.lbs[[4]])`, $t$(`r sprintf("%i", cor.test.inches.lbs[[2]])`)=`r sprintf("%2.3f", cor.test.inches.lbs[[1]])`, $p$=`r sprintf("%.5f", cor.test.inches.lbs[[3]])`, with a 95$\%$ confidence interval of [`r sprintf("%.3f", cor.test.inches.lbs[[9]][1])`, `r sprintf("%.3f", cor.test.inches.lbs[[9]][2])`].

<span class="blue">I did some formatting to make the numbers look pretty.</span>

`sprintf("%.3f", my.var)` limits `my.var` to 3 decimal places; where `sprintf("%2.3f", my.var)` limits it to 2 digits to the left and 3 to the right.

## Now back to analysis with our synthetic data

```{r linear-reg}
fit <- lm(formula = inches ~ lbs, data = height.weight)
summary(fit) # Use lm() command to fit formula
(ci <- confint(fit)) # confint() command fits confidence intervals
```

Surrounding `(ci <- confint(fit))` in parentheses saves variable ci and prints it out.

## How'd we do?

| Parameter | Actual    | Low Estimate | High Estimate |
|-----------|-----------|--------------|---------------|
| $\beta0$  | `r beta0` | `r ci[1,1]`  | `r ci[1,2]`   |
| $\beta1$  | `r beta1` | `r ci[2,1]`  | `r ci[2,2]`   |

- Why off on the slope ($\beta1$)
- Random error, probably. Could run again.

---

```{r run-it-all-again, echo=FALSE}
# random weights between 80 lbs and 250 lbs (uniform sampling)
w <- runif(n = sample.n, min = 80, max = 250)

h.pred <- rep(x = beta0, n = sample.n) + beta1 * w
h <- h.pred + rnorm(n = sample.n, mean = 0, sd = sigma)
height.weight <- data.frame(inches = h, lbs = w)

fit <- lm(formula = inches ~ lbs, data = height.weight)
summary(fit)
(ci <- confint(fit)) # saves in variable ci and prints
```

---

| Parameter | Actual    | Low Estimate | High Estimate |
|-----------|-----------|--------------|---------------|
| $\beta0$  | `r beta0` | `r ci[1,1]`  | `r ci[1,2]`   |
| $\beta1$  | `r beta1` | `r ci[2,1]`  | `r ci[2,2]`   |

## Simulation of fMRI data

- Critical review: Welvaert, M., & Rosseel, Y. (2014). A Review of fMRI Simulation Studies. PLOS ONE, 9(7), e101953. <https://doi.org/10.1371/journal.pone.0101953>.
- Welvaert, M., Durnez, J., Moerkerke, B., Berdoolaege, G. & Rosseel, Y. (2011). neuRosim: An R Package for Generating fMRI Data. Journal of Statistical Software, 44(10). Retrieved from <https://www.jstatsoft.org/article/view/v044i10>
- AFNI's *AlphaSim*, <https://afni.nimh.nih.gov/pub/dist/doc/program_help/AlphaSim.html>

# Visualization in R

## Plot first, analyze last

>- Why?
  + [Mike Meyer](https://www.linkedin.com/in/mike-meyer-0056578a) told me so
  + Less biased
  + Easier to be transparent and reproducible
  + Want/need to plot eventually anyway
  + If a picture's worth a thousand words...
>- How?

## How

- Base graphics
    + `plot(x,y)` `hist(x)`, `coplot()`
- `ggplot2`
    + Grammar of graphics
    
## Base graphics

- Try it, maybe you'll like it
- `plot()` takes many types of input
- So does `summary()`
- A little harder to customize

## Data visualization with ggplot2

Wickham, H. & Grolemund, G. (2017). *R for Data Science*. O'Reilly. <http://r4ds.had.co.nz/>

## Let's just walk through the data visualiation chapter

<http://r4ds.had.co.nz/data-visualisation.html>

## Other ggplot2 resources

- Wickham, H. (2010). *ggplot2: Elegant Graphics for Data Analysis (Use R!)* <http://ggplot2.org/book/>
- ggplot2 2.1.0 documentation: <http://docs.ggplot2.org/current/>
